### 2018-01-30 includes/abstract.tex
adding comma
editing first pass
cleaned up the paper a lot

\begin{abstract}

    In this paper, we introduce \fuzzer{}: a modular \smt{} problem transformer and generator for string solvers. We supply a repository of problems generated by \fuzzer{} in \smtfull{} format. We systemically compare \theSolvers{} on groups of such problems focusing on different combinations of string operators, and identify problems that are particularly challenging for some solvers. We briefly explain our observations and showcase how \fuzzer{} helped identify causes of the performance degradations in \us{}.

\end{abstract}

>>>>

\begin{abstract}
    In this paper, we introduce \fuzzer{}: a modular \smt{} problem
    transformer and generator for string solvers. We supply a
    repository of problems generated by \fuzzer{} in \smtfull{}
    format. We systemically compare \theSolvers{} on groups of such
    problems focusing on different combinations of string operators,
    and identify problems that are particularly challenging for some
    solvers. We briefly explain our observations and showcase how
    \fuzzer{} helped identify causes of performance degradations in
    \us{}.
\end{abstract}

### 2018-01-30 includes/experiment.tex
cleaned up the paper a lot

\section{Usefulness to \us{}: A Case Study}
\vspace{-0.1in}

        We describe the defects and missing heuristics exposed by \fuzzer{} to illustrate why \fuzzer{} is a valuable asset in solver development.

\vspace{-0.15in}
\subsection{Defects discovered and fixed in \us{}}
\vspace{-0.1in}

The first issue we found caused \us{} to perform poorly when equivalence classes 
of arithmetic terms were very large (e.g. the lengths of many strings were equal
to 0). This issue was detected by the \texttt{Concats} problem suite. The slowdown 
was caused by a loop over terms in the same equivalence class to search for an 
integer constant. The fix was to check the root of the equivalence class for the 
integer constant, as Z3 makes ``interpreted terms'' (e.g. constants) the root of 
equivalence classes whenever possible. This issue was last observed in commit 
\texttt{6308636}, and was fixed in commit \texttt{3865c45}.

The second issue, which was exposed by the \texttt{Lengths} suite, caused \us{} to 
perform poorly when a variable has a large length constraint.
The slowdown was caused by the solver performing a search for a satisfying length
even when the input formula explicitly constrained the exact length of a string term.
The fix was to check whether the integer solver had already assigned a length value
before searching for length, and skip the search if this was the case.
This issue was last observed in commit \texttt{66bc68f}, and was fixed in commit 
\texttt{7b536e9}.    
%The slowdown was  caused by the solver performing a linear search for a satisfying length when 
%producing a model for the variable. This issue was an inconsistency between the 
%implementation and the behavior described in the \us{} paper \cite{z3str3}. 
%This issue was last observed in commit \texttt{66bc68f}, and was fixed in commit 
%\texttt{7b536e9}.

\vspace{-0.15in}    
\subsection{Solving heuristics}
\vspace{-0.1in}

In order to improve the efficiency of any solver, discovering new heuristics is crucial.
As an illustrative example, consider detection of conflicting constraints.
A solver that follows the string constraint solving algorithm and gradually propagates
the (dis)equalities discovered along the way will eventually detect a conflict.
However, in many scenarios, a conflict can be discovered ``early'' before trying to
figure out the relationships among subterms and propagating the facts.
The benefits of this eager approach are clear: this saves the effort of reasoning about
subterms and propagating observations, as a conflict will eventually be reached in either case,
forcing conflict analysis and a backjump.
As reported in existing works, heuristics of this form usually lead to significant improvements.
However, the process of discovering such heuristics is extremely challenging,
as they are usually learned from a class of examples whose commonalities
are not always obvious or immediately indicative of such inefficiencies.

The instances in the \texttt{concats-big} suite generated by \fuzzer{} helped us 
discover a missing heuristic in \us{}. In particular, \us{} didn't make full use 
of the solving context (e.g. some terms are empty strings) to simplify the 
concatenations of a long list of string terms before trying to reason about the 
equivalences among subterms. \us{} therefore introduced a large number of unnecessary
intermediate variables and propagations.
As an immediate consequence of the \texttt{concats-big} benchmarks, we were able to identify
the factors common to these instances and rapidly formulate a hypothesis as to why they were
performing poorly. This allowed us to move forward with the development of new heuristics
to resolve these issues and achieve better performance.

% In this section, we analyse our experimental results. We discuss the causes of 
% \us{}'s poor performance on the two problem suites: \textit{concats-balanced} 
% and \textit{concats-big}. We were not equipped well enough to debug the 
% behaviours of the other solvers, but we provide our data in hopes that they will 
% be useful to the solvers' authors.

%     \subsection{\us{} on concats-balanced}
%         \todo{Dmitry with help: Explain why \us{} performs poorly on this suite.}

% X = "solution"
% X = (((A.B).(C.D)).((E.F).(G.H))).
%     (((I.J).(K.L)).((M.N).(O.P)))

    % \subsection{\us{} on concats-big}

    %     \todo{Dmitry with help: Explain why \us{} performs poorly on this suite.}

% X = "solution"
% X = (A.(B.(C.(D.(E.(F.(G.H)))))))

>>>>

\section{Usefulness to \us{}: A Case Study}
\label{sec:analysis}

We found a number of performance-related issues and opportunities for
new heuristics in \us{} thanks to \fuzzer{}. For example, the
instances in the \texttt{concats-big} suite generated by \fuzzer{}
helped us discover a missing heuristic. In particular, \us{} didn't
make full use of the solving context (e.g. some terms are empty
strings) to simplify the concatenations of a long list of string terms
before trying to reason about the equivalences among subterms. \us{}
therefore introduced a large number of unnecessary intermediate
variables and propagations.  As an immediate consequence of the
testing \us{} on \texttt{concats-big} benchmarks, we were able to
identify the factors common to these instances and rapidly formulate a
hypothesis as to why they were performing poorly. This allowed us to
move forward with the development of new heuristic to resolve this
issue and achieve better performance.

%% %VG: this subsection is not needed in the paper    
%% \noindent{\textbf{Defects discovered and fixed in \us{}:}} The first
%% issue we found caused \us{} to perform poorly when equivalence classes
%% of arithmetic terms were very large (e.g. the lengths of many strings
%% were equal to 0). This issue was detected by the \texttt{Concats}
%% problem suite. The slowdown was caused by a loop over terms in the
%% same equivalence class to search for an integer constant. The fix was
%% to check the root of the equivalence class for the integer constant,
%% as Z3 makes ``interpreted terms'' (e.g. constants) the root of
%% equivalence classes whenever possible. The second issue, which was
%% exposed by the \texttt{Lengths} suite, caused \us{} to perform poorly
%% when a variable has a large length constraint.  The slowdown was
%% caused by the solver performing a search for a satisfying length even
%% when the input formula explicitly constrained the exact length of a
%% string term.  The fix was to check whether the integer solver had
%% already assigned a length value before searching for length, and skip
%% the search if this was the case.

%% In order to improve the efficiency of any solver, discovering new
%% heuristics is crucial.  As an illustrative example, consider the
%% problem of how the a string solver may detect conflicting
%% constraints. A string solver that gradually propagates
%% (dis)equalities discovered along the way will eventually detect a
%% conflict.  However, in many scenarios, a conflict can be discovered
%% ``early'' before trying to figure out the relationships among
%% subterms and propagating the facts.  The benefits of this eager
%% approach are clear: this saves the effort of reasoning about
%% subterms and propagating observations, as a conflict will
%% eventually be reached in either case, forcing conflict analysis and
%% a backjump.  As reported in existing works, heuristics of this form
%% usually lead to significant improvements.  However, the process of
%% discovering such heuristics is extremely challenging, as they are
%% usually learned from a class of examples whose commonalities are
%% not always obvious or immediately indicative of such
%% inefficiencies.

### 2018-01-30 includes/introduction.tex
cleaned up the paper a lot

    Many SMT solvers supporting the theory of strings (e.g., CVC4 \cite{cvc4},  Z3str2 \cite{z3str2}, Norn \cite{norn} and Z3str3 \cite{z3str3}) have been developed over the past decade. To validate and benchmark their solvers, many authors have relied on hand-crafted input suites~\cite{cvc4-tests,z3str3-tests,z3str2-tests} or real-world examples from industrial applications~\cite{kaluza,kausler}. These test suites have helped developers identify implementation defects and develop more sophisticated solving heuristics. Although extremely helpful, these suites can only cover a limited range of the problem space because they are either manually created or translated from carefully developed programs.

    Fuzzing, on the other hand, has already been widely used in similar scenarios to systematically generate more diverse test cases. Guided by different rules, it can also be applied to generate different instances that are particularly designed to challenge selected parts of the solvers. Inspired by traditional fuzzers, we introduce a modular fuzzer for string solvers, \fuzzer{}, and describe its utility as an exploratory validation tool. We demonstrate its effectiveness by presenting defects and limitations we were able to identify in leading solvers.
%     Moreover, a standard input language for problem instances, \smtfull{}\cite{smt}, has been developed and widely adopted.

    \subsection{Contributions}

        \begin{enumerate}
            \item \textbf{The \fuzzer{} tool}: We describe in Section~\ref{sec:fuzzer} a modular fuzzer which can transform and generate \smtfull{} string and regex problems. We briefly document its components and modular architecture. We provide example use cases to demonstrate its usefulness as an exploratory tool for validating solvers. We additionally detail its regex generation and transformation capabilities.
            \item \textbf{A repository of \smtfull{} problems}: We present a repository of \smtfull{} string and regex problem suites we generated using \fuzzer{} in Section~\ref{sec:problems}. We demonstrate the usefulness of these suites by revealing specific defects they helped uncover and fix in \us{}.
            \item \textbf{Experimental Results}: We compare \theSolvers{} on the \fuzzer{} suites and selectively present the experimental result on \theSuites{} in Section~\ref{sec:data}. We highlighted these suites because they uniquely made some solvers perform poorly, but not others.
            \item \textbf{Analysis}: We analyze our experimental results in Section~\ref{sec:analysis} by inspecting the execution traces of the solvers on the given problems. We pinpoint algorithmic limitations in \us{} that cause poor performance, and analyze proposed and implemented enhancements that address the issues which \fuzzer{} helped us to identify.
        \end{enumerate}

>>>>

In recent years many algorithms for solving string constraints have
been developed and implemented as part of SMT solvers such as CVC4
\cite{cvc4} and Z3 (e.g., Z3str2 \cite{z3str2} and Z3str3
\cite{z3str3}). To validate and benchmark these solvers, their
developers have relied on hand-crafted input
suites~\cite{cvc4-tests,z3str3-tests,z3str2-tests} or real-world
examples from a limited class of industrial
applications~\cite{kaluza,kausler}. These test suites have helped
developers identify implementation defects and develop more
sophisticated solving heuristics. Although extremely helpful, these
suites are of limited value since they do not test the vast majority
of interesting behaviors in solvers. There is an acute need for more
robst, inexpensive, and automatic way of generating test suites to
test SMT solvers, that can augment hand-crafted as well as application
tests.

Random and mutation fuzzing are widely used test all kinds of software
including SAT solvers~\cite{}. Inspired by their utility, we introduce
a fuzzer for string solvers, \fuzzer{}, and describe its value as an
exploratory validation tool. We demonstrate the efficacy \fuzzer{} by
presenting defects and limitations it enabled us to identify in
leading string solvers, which otherwise would have been difficult to
detect. To the best of our knowledge, \fuzzer{} is the first such tool
aimed at automatic generation of string constraints. \fuzzer{} allows
to test string solvers by enabling us to mutate existing classes of
benchmarks in interesting ways, as well as generate extremal inputs
(e.g., inputs with very long strings or deep concatenations or
highly-complex mix of string operations and regular
expressions). Extremal inputs are particularly useful in empirically
understanding asymptotic behavior of solvers in a way that no other
class of instances can.

\noindent{\textbf{Contributions:}}

\begin{enumerate}
\item \textbf{The \fuzzer{} tool}: We describe in
  Section~\ref{sec:fuzzer} a modular fuzzer which can transform and
  randomly generate extremal \smtfull{} string and regex problems. We
  briefly document its components and modular architecture. We provide
  example use cases to demonstrate its utility as an exploratory tool
  for validating solvers.

\item \textbf{A repository of \smtfull{} problems}: We present a
  repository of \smtfull{} string and regex problem suites we
  generated using \fuzzer{} in Section~\ref{sec:problems}. We
  demonstrate the usefulness of these suites by revealing specific
  defects they helped uncover and fix in \us{}.

\item \textbf{Experimental Results}: We compare \theSolvers{} on the
  \fuzzer{} suites and selectively present the experimental result on
  \theSuites{} in Section~\ref{sec:data}. We highlighted these suites
  because they uniquely made some solvers perform poorly, but not
  others.

\item \textbf{Analysis}: We analyze our experimental results in
  Section~\ref{sec:analysis} by inspecting the execution traces of the
  solvers on the given problems. We pinpoint algorithmic limitations
  in \us{} that cause poor performance, and analyze proposed and
  implemented enhancements that address the issues which \fuzzer{}
  helped us to identify.
\end{enumerate}

### 2018-01-30 includes/problems.tex
cleaned up the paper a lot

\section{Problem Suites}
\label{sec:problems}

In this section, we describe the benchmark suites that we generated with \fuzzer{} on which the experimental evaluation was conducted.

\subsection{Generated suite}

    We generated several suites of string and regex problems using \fuzzer{}\footnote{For review anonymity, we removed the link to the source code and the generated suites. They will be provided separately in the artifact submission.}. Table~\ref{tbl:generated} describes the suite \texttt{generated} by \generator{}.

@@

\subsection{Transformed suite}

    The \texttt{transformed} suite contains benchmarks derived from existing seed problems by iteratively applying \transformer{}. Table~\ref{tbl:transformed} shows the overview. Every problem in the suite is named according to its seed and the transformations it undertook. For example, input \texttt{seed-fuzz-graft.smt2} was transformed by applying \texttt{fuzz} and then \texttt{graft} from \texttt{seed.smt2}.

@@

    The Amazon category is obtained from two seeds: (1) \texttt{z3-regex-1.smt2} encodes the assertions \texttt{(S not in aaabbb*)} and \texttt{(S in aaa*bbb*)} for a string constant S; (2) \texttt{z3-regex-2-negative.smt2} includes regex \texttt{(S in aaabbb*)} and \texttt{(S not in aaa*bbb*)}.
    Both problems come from our industrial collaborators. In total, 516 unique problems were generated from these two seeds.

    The Regex category is seeded by the z3str2 regular expression test suite \cite{z3str2-tests}.
    This test suite contains 42 problems dealing with a wide variety of regular expressions.
    Through cumulative transformations, we have $10,577$ unique benchmarks.

>>>>

\section{Problem Suites}
\label{sec:problems}
In this section, we describe the benchmark suites that we generated
with \fuzzer{} on which the experimental evaluation was conducted.

\noindent{\textbf{Generated Benchmarks:}} We generated several suites
of string and regex problems using \fuzzer{}\footnote{For review
  anonymity, we removed the link to the source code and the generated
  suites. They will be provided separately in the artifact
  submission.}. Table~\ref{tbl:generated} describes the suite
\texttt{generated} by \generator{}.

@@

\noindent{\textbf{Transformed Benchmarks:}} The \texttt{transformed}
suite contains benchmarks derived from existing seed problems by
iteratively applying \transformer{}. Table~\ref{tbl:transformed} shows
the overview. Every problem in the suite is named according to its
seed and the transformations it undertook. For example, input
\texttt{seed-fuzz-graft.smt2} was transformed by applying
\texttt{fuzz} and then \texttt{graft} from \texttt{seed.smt2}.

@@

The Amazon category is obtained from two seeds: (1)
\texttt{z3-regex-1.smt2} encodes the assertions \texttt{(S not in
  aaabbb*)} and \texttt{(S in aaa*bbb*)} for a string constant S; (2)
\texttt{z3-regex-2-negative.smt2} includes regex \texttt{(S in
  aaabbb*)} and \texttt{(S not in aaa*bbb*)}.  Both problems come from
our industrial collaborators. In total, 516 unique problems were
generated from these two seeds. The Regex category is seeded by the
z3str2 regular expression test suite \cite{z3str2-tests}.  This test
suite contains 42 problems dealing with a wide variety of regular
expressions.  Through cumulative transformations, we have $10,577$
unique benchmarks.

### 2018-01-30 includes/stringfuzz.tex
add section labels; remove manual spacing; fix sectioning
cleaned up the paper a lot

\section{\fuzzer{}}
\label{sec:fuzzer}
\vspace{-0.1in}

%     This section describes the fuzzer \fuzzer{}, which generates new \smtfull{} problems and transforms existing ones. 
    \fuzzer{} is implemented as a Python package, and comes with several executables to generate, transform, and analyze \smtfull{} string problems. It can either be installed from source, or from the Python PIP package repository.

%     {\it For review anonymity, we removed the link to the source code. It will be provided separately in the artifact submission.}

    \fuzzer{} is organized to be easily extended: all generators and transformers are independent Python modules. Its components are also implemented as \unix{} ``filters'', to enable easy integration with other tools (including themselves). For example, the outputs of generators can be piped into transformers, and transformers can be chained to produce a stream of tuned inputs to a solver.

    \fuzzer{} is composed of the following tools:

\vspace{-0.15in}
\subsubsection{\generator{}:}
% \vspace{-0.1in}

This tool generates \smt{} problems. It supports several generators and options that specify its output. Details can be found in Table~\ref{tbl:generators}.

\begin{table}[t]
\vspace{-0.25in}
\centering
\caption{\generator{} generators.}
\label{tbl:generators}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
                  \begin{tabular}{|l|l|}
                    \hline
                    \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
                    \textit{concats}     & Long concats and optional random extracts. \\ \hline
                    \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
                    \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
                    \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
                    \textit{regex}       & Regexes of varying complexity. \\ \hline
                    \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
                    \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
                \end{tabular}
  }
  \egroup
\vspace{-0.2in}
}
\end{table}

\vspace{-0.15in}
\subsubsection{\transformer{}:}
% \vspace{-0.1in}

This tool transforms \smt{} problems. It supports several transformers and options that specify its output and input, which are explained in Table~\ref{tbl:transformers}. Note that transformers \textit{translate} and \textit{reverse} also preserve problem satisfiability \cite{ifaz}.

\begin{table}[t]
\vspace{-0.25in}
\centering
\caption{\transformer{} transformers}
\label{tbl:transformers}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
                \begin{tabular}{|l|l|}
                    \hline
                    \textbf{Name}        & \textbf{The transformer ...} \\ \hline
                    \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
                    \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
                    \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
                    \textit{nop}         & Does nothing. \\ \hline
                    \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
                    \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
                    \textit{translate}   & Maps every character X to a different character Y. \\ \hline
                    \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
                \end{tabular}
  }
  \egroup
  \vspace{-0.2in}
}
\end{table}

\vspace{-0.15in}
\subsubsection{\texttt{stringstats}:}

This tool takes an \smt{} problem as input and outputs its properties: the number of variables/literals, the max/median syntactic depth of expressions, the max/median literal length, etc.

% \texttt{stringstats} is handy when analyzing a problem.

%     \begin{description}

@@

    \section{Regex Fuzzing Capabilities}
    \vspace{-0.1in}
        \fuzzer{} can generate and transform problems with regular expression constraints. The command \texttt{stringfuzzg regex} invokes the regex problem generator. It generates a problem of the form: %\footnote{Variable declarations omitted for brevity.}:
        \begingroup
        \footnotesize
        \begin{align*}
            & \texttt{(assert (str.in.re X}\; R_0\; \texttt{))} \\
            & \texttt{(assert (str.in.re X}\; R_n\; \texttt{))}* \\
            & \texttt{(assert (<= Min (str.len X)))}? \\
            & \texttt{(assert (<= (str.len X)) Max)}?
        \end{align*}
        \endgroup
        where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the problem is a set of one or more regex constraints on a single variable, with optional maximum and minimum length constraints. The regex constraints $R$ are each of the form:
        \begingroup
        \footnotesize
        \begin{align*}
            & \texttt{(re.++}\; \ T_0\; \  T_1\; \texttt{...}\; T_n\; \texttt{)}
        \end{align*}
        \endgroup
        and each $T_i$ is a recursive term of the form:
        \begingroup
        \footnotesize
        \begin{align*}
            & \texttt{(re.*}\; T_{i_j}\; \texttt{) | (re.+}\; T_{i_j}\; \texttt{) | (re.union}\; T_{i_{j_1}}\; T_{i_{j_2}}\; \texttt{)}
        \end{align*}
        \endgroup
        where $j$  is the specified depth of recursion. Terms at depth 0 are regex constants. Informally, this form describes a concatenation of regex terms. Each term is a random nested regex operator (chosen from regex Kleene star, repetition, and union), up to a specified depth, terminating in a regex literal. Below are several example regexes (with spacing added for clarity) of depth 2 produced by this scheme:
        \begingroup
        \footnotesize
        \begin{align*}
            & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad ((\texttt{ddd})*)+\quad ((\texttt{ee})+|(\texttt{fff})*)
        \end{align*}
        \endgroup

\section{Example Use Cases}
\vspace{-0.1in}

        \fuzzer{} can be used to generate problems to be saved as static suites. It can also be used in a randomised fashion for exploratory debugging. As an example, \fuzzer{} can be invoked as follows to feed random problems to \cvc{} until it returns an error:
\begin{center}
\vspace{-0.05in}
\begin{varwidth}{\linewidth}
{\scriptsize\begin{verbatim}while stringfuzzg -r random | tee problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do
    sleep 0
done
\end{verbatim}}
\end{varwidth}
\end{center}
        Some \unix{} scripting can yield other interesting results. For example, the following script finds a regex problem on which \us{} times out after 5 seconds:
\begin{center}
\vspace{-0.1in}
\begin{varwidth}{\linewidth}
        {\scriptsize\begin{verbatim}while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > problem.smt25
    result=`z3str3 -T:5 problem.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat problem.smt25
        break
    fi
done\end{verbatim}}
\end{varwidth}
\end{center}

>>>>

\section{\fuzzer{}}
\label{sec:fuzzer}
\fuzzer{} is implemented as a Python package, and comes with several
executables to generate, transform, and analyze \smtfull{} string
problems. It can either be installed from source, or from the Python
PIP package repository. \fuzzer{} is organized to be easily extended:
all generators and transformers are independent Python modules. Its
components are also implemented as \unix{} ``filters'', to enable easy
integration with other tools (including themselves). For example, the
outputs of generators can be piped into transformers, and transformers
can be chained to produce a stream of tuned inputs to a
solver.

\noindent{Specifically, \fuzzer{} is composed of the following tools:}

\noindent{\textbf{\generator{}:}} This tool generates \smt{}
problems. It supports several generators and options that specify its
output. Details can be found in Table~\ref{tbl:generators}.

\begin{table}[t]
%\vspace{-0.25in}
\centering
\caption{\generator{} generators.}
\label{tbl:generators}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
    \begin{tabular}{|l|l|}
      \hline
      \textbf{Name}        & \textbf{Generated Problems have ...} \\ \hline
      \textit{concats}     & Long concats and optional random extracts. \\ \hline
      \textit{lengths}     & Many variables (and their concats) with length constraints. \\ \hline
      \textit{overlaps}    & An expression of the form A.X = X.B. \\ \hline
      \textit{equality}    & An equality among concats, each with variables or constants. \\ \hline
      \textit{regex}       & Regexes of varying complexity. \\ \hline
      \textit{random-text} & Random, likely syntactically \textit{in}valid text. \\ \hline
      \textit{random-ast}  & Random, but semantically \textit{valid} text. \\ \hline
    \end{tabular}
  }
  \egroup
%\vspace{-0.2in}
}
\end{table}

\noindent{\textbf{\transformer{}:}This tool transforms \smt{}
  problems. It supports several transformers and options that specify
  its output and input, which are explained in
  Table~\ref{tbl:transformers}. Note that transformers
  \textit{translate} and \textit{reverse} also preserve problem
  satisfiability \cite{ifaz}.

\begin{table}[t]
%\vspace{-0.25in}
\centering
\caption{\transformer{} transformers}
\label{tbl:transformers}
{
  \footnotesize
  \bgroup
  \def\arraystretch{1}
  \resizebox{0.8\columnwidth}{!}{
    \begin{tabular}{|l|l|}
      \hline
      \textbf{Name}        & \textbf{The transformer ...} \\ \hline
      \textit{fuzz}        & Randomly replaces literals and operators with similar ones. \\ \hline
      \textit{graft}       & For each sort, swaps a random non-leaf node with a leaf node. \\ \hline
      \textit{multiply}    & Multiplies each number by N, and repeats each character N times. \\ \hline
      \textit{nop}         & Does nothing. \\ \hline
      \textit{reverse}     & Reverses all string literals and concat arguments. \\ \hline
      \textit{rotate}      & Performs a random syntax tree rotation on two concats. \\ \hline
      \textit{translate}   & Maps every character X to a different character Y. \\ \hline
      \textit{unprintable} & Randomly replaces all characters in literals with unprintable ones. \\ \hline
                \end{tabular}
  }
  \egroup
  %\vspace{-0.2in}
}
\end{table}

\noindent{\textbf{\texttt{stringstats}:}} This tool takes an \smt{}
problem as input and outputs its properties: the number of
variables/literals, the max/median syntactic depth of expressions, the
max/median literal length, etc.

\noindent{\textbf{Regex Fuzzing Capabilities:}} \fuzzer{} can generate
and transform problems with regular expression constraints. The
command \texttt{stringfuzzg regex} invokes the regex problem
generator. It generates a problem of the form:

\begingroup
\footnotesize
\begin{align*}
  & \texttt{(assert (str.in.re X}\; R_0\; \texttt{))} \\
  & \texttt{(assert (str.in.re X}\; R_n\; \texttt{))}* \\
  & \texttt{(assert (<= Min (str.len X)))}? \\
  & \texttt{(assert (<= (str.len X)) Max)}?
\end{align*}
\endgroup

where $R_i \in RegEx$, and $Min, Max \in Int$. More simply, the
problem is a set of one or more regex constraints on a single
variable, with optional maximum and minimum length constraints. The
regex constraints $R$ are each of the form:

\begingroup
\footnotesize
\begin{align*}
  & \texttt{(re.++}\; \ T_0\; \ T_1\; \texttt{...}\; T_n\; \texttt{)}
\end{align*}
\endgroup
and each $T_i$ is a recursive term of the form:
\begingroup
\footnotesize
\begin{align*}
  & \texttt{(re.*}\; T_{i_j}\; \texttt{) | (re.+}\; T_{i_j}\;
  \texttt{) | (re.union}\; T_{i_{j_1}}\; T_{i_{j_2}}\; \texttt{)}
\end{align*}
\endgroup

where $j$ is the specified depth of recursion. Terms at depth 0 are
regex constants. Informally, this form describes a concatenation of
regex terms. Each term is a random nested regex operator (chosen from
regex Kleene star, repetition, and union), up to a specified depth,
terminating in a regex literal. Below are several example regexes
(with spacing added for clarity) of depth 2 produced by this scheme:

\begingroup
\footnotesize
\begin{align*}
  & ((\texttt{a}|\texttt{b})|(\texttt{cc})+)\quad
  ((\texttt{ddd})*)+\quad ((\texttt{ee})+|(\texttt{fff})*)
\end{align*}
\endgroup

\noindent{\textbf{Example Use Cases:}} \fuzzer{} can be used to
generate benchmark suites in a batch mode, or can be used in an online
randomized fashion for exploratory debugging. For example, \fuzzer{}
can be invoked as follows to feed random problems to \cvc{} until it
returns an error:

\begin{center}
%\vspace{-0.05in}
\begin{varwidth}{\linewidth}
{\scriptsize\begin{verbatim}while stringfuzzg -r random | tee
  problem.smt25 | cvc4 --lang smt2.5 --tlimit=5000; do sleep 0 done
\end{verbatim}}
\end{varwidth}
\end{center}

Some \unix{} scripting can yield other interesting results. For
example, the following script finds a regex problem on which \us{}
times out after 5 seconds:

\begin{center}
%\vspace{-0.1in}
\begin{varwidth}{\linewidth}
        {\scriptsize\begin{verbatim}while true; do
    stringfuzzg -r regex -c -a increasing -r 2 -t 5 -d 3 > problem.smt25
    result=`z3str3 -T:5 problem.smt25`
    echo $result
    if [ "$result" == "timeout" ]; then
        cat problem.smt25
        break
    fi
done\end{verbatim}}
\end{varwidth}
\end{center}

% \texttt{stringstats} is handy when analyzing a problem.    
%     \begin{description}

@@
